---
title: Literate Analysis with Quarto
editor: 
  markdown: 
    wrap: 72
---

::: callout-tip
## Learning Objectives

After completing this session, you will be able to:

-   Describe the concept of literate analysis and how it supports
    reproducible data science
-   Format prose (non-code) text in a Quarto document using Markdown
    syntax
-   Incorporate functional code chunks into a Quarto document and
    customize their output
-   Build and render a complete example analysis
:::

## Introduction to Literate Programming

Traditional code is written for computers - it can be hard for humans to
follow, even the person who wrote it. In 1984, computer science pioneer
Donald Knuth proposed flipping this around: rather than instructing a
computer, write code that *explains your thinking to other people*.

> "Instead of imagining that our main task is to instruct a computer
> what to do, let us concentrate rather on explaining to human beings
> what we want a computer to do." - Donald Knuth, 1984

In a **literate analysis**, we weave together natural-language prose
(describing your goals, methods, and interpretations) with functional
code (that actually performs the analysis). The result renders into a
polished HTML, PDF, or Word report - a single document that is both
human-readable and fully reproducible.

<!-- All too often, computational methods are written in such a way as to be borderline incomprehensible even to the person who originally wrote the code! The reason for this is obvious: computers interpret information very differently than people do. In 1984, Donald Knuth proposed a reversal of the programming paradigm by introducing the concept of *Literate Programming* ([Knuth 1984](http://www.literateprogramming.com/knuthweb.pdf)). -->

<!-- > *"Instead of imagining that our main task is to instruct a computer what to do, let us concentrate rather on explaining to human beings what we want a computer to do."* -->

<!-- If our aim is to make scientific research more transparent, the appeal of this paradigm reversal is immediately apparent. By switching to a literate analysis model, **you help enable human understanding of what the computer is doing**. As Knuth describes, in the literate analysis model, the author is an "*essayist*" who chooses variable names carefully, explains what they mean, and introduces concepts in the analysis in a way that facilitates understanding. -->

<!-- In practice, literate programming is achieved by interleaving sections of natural language prose that describe the goals, methods, and interpretation of an analysis, with sections of code that perform the analysis, generate figures, and save results.  The resulting document is usually rendered as a nicely-formatted report in HTML, MS Word, or PDF format.  Contrast this with a more traditional script, which includes only functional code with (hopefully) a few brief explanatory comments. -->

## Quarto and Literate Programming

**Quarto** is an open-source publishing system that makes literate
analysis straightforward. A Quarto file (`.qmd`) mixes formatted prose
with executable **code chunks**, and renders them together into a
complete report.

::: callout-tip
### Coding languages in Quarto

We will focus on using Quarto with R code, but it is designed to work
seamlessly with other coding languages as well, including Python, Julia,
and Observable JavaScript. You can even use multiple languages in the
same document.
:::

::: callout-note
### Quarto vs. RMarkdown

If you've used RMarkdown (`.Rmd`) before, Quarto (`.qmd`) will feel
familiar. The main differences are that Quarto has richer multi-language
support and improved formatting options. Both formats share the same
three-part structure: YAML, prose, and code chunks.
:::

### Why does this matter for reproducibility?

When your entire analysis - data loading, cleaning, statistics, and
figures - lives in one Quarto document, you can re-run everything with a
single click. If the data updates, so does your report. This eliminates
error-prone copy-pasting between scripts, spreadsheets, and documents,
and gives collaborators (and your future self!) a clear trail from raw
data to final results.

<!-- In an R script, the language assumes that you are writing R code, unless you specify that you are writing prose (using a comment, designated by `#`). The paradigm shift of literate analysis comes in the switch to RMarkdown or Quarto, where instead of assuming you are writing code, they assume that you are writing prose unless you specify that you are writing code. This, along with the formatting provided by markdown, encourages the "essayist" to write understandable prose to accompany the code that explains to the human-beings reading the document what the author told the computer to do. This is in contrast to writing just R code, where the author telling to the computer what to do with maybe a smattering of terse comments explaining the code to a reader. -->

### Example: Literate Analysis From Quarto

Let's look at an example of what a rendered literate analysis looks like
using a real example. [Here is an
example](https://nceas.github.io/sasap-training/materials/reproducible_research_in_r_fairbanks/example-brood-table-analysis.html)
of an analysis workflow written using RMarkdown. Note that if this
analysis would be in Quarto, the render version it would be similar,
except for formatting and layout (eg: the default font in Quarto is
different).

There are a few things to notice about this document, which assembles a
set of similar data sources on salmon brood tables with different
formatting into a single data source.

-   It introduces the data sources using in-line images, links,
    interactive tables, and interactive maps.
-   An example of data formatting from one source using R is shown.
-   The document executes a set of formatting scripts in a directory to
    generate a single merged file.
-   Some simple quality checks are performed (and their output shown) on
    the merged data.
-   Simple analysis and plots are shown.

In addition to achieving literate analysis, this document also
represents a **reproducible analysis**. Because the entire merging and
quality control of the data is done using the R code in the Quarto file,
if a new data source and formatting script are added, the document can
be run all at once with a single click to re-generate the quality
control, plots, and analysis of the updated data.

## Structure of a Quarto Document

Every Quarto document has three components:

-   **YAML header** - metadata at the very top of the file, enclosed in
    `---`. It controls the document title, author, output format, and a
    lot more.
    
:::{.callout-tip appearance="simple" icon=false}

``` yaml
---
title: "My Analysis"
author: "Your Name"
format: html
---
```
:::

-   **Prose** - plain text written in Markdown (the white areas of the
    document).
    
:::{.callout-tip appearance="simple" icon=false}
:::

-   **Code chunks** - executable code blocks (the grey areas), opened
    with ```` ```{r} ```` and closed with ```` ``` ````.
    
:::{.callout-tip appearance="simple" icon=false}

```{{r}}
x <- "Hello world!"
y <- 1 + 1
```

<!--NOTE: the double curly braces mean this will *display* as an R code chunk including header and footer, but not execute -->
:::

## Let's Create A Quarto Document

:::{.panel-tabset}

#### RStudio

1.  In RStudio, go to **File** {{< fa arrow-right >}} **New File**
    {{< fa arrow-right >}} **Quarto Document**.
2.  Give it a title (e.g., "Introduction to Quarto"), leave the other
    options as-is, then click **Create**.
3.  Look at what you get: a YAML header at the top, with title and format, and a blank canvas below.
    
#### Positron

1.  In Positron, go to **File** {{< fa arrow-right >}} **New File**
    {{< fa arrow-right >}} **Quarto Document**.
2.  Look at what you get: a YAML header at the top for title and format, and a blank canvas below.
3.  Fill in the title (e.g., "Introduction to Quarto")

:::

Below the YAML header, let's add some content.  Type or copy/paste so your document looks like this:

:::{.callout-note collapse="true"}

### Expand this block and add this code to your Quarto document

```yaml
---
title: "Untitled"
format: html
---
```
```
## Quarto

### Markdown section
In this section, include prose text formatted using Markdown syntax.

### Code chunks

You can also include code chunks:
```
```{{r basic math}}
x <- 1 + 1
x
```
```
## Including Plots

You can also embed plots, for example:
```
```{{r pressure plot}}
#| echo: false
plot(pressure)
```
:::

Click the **Render** button. You will be asked to save first - give it a useful name (e.g., `quarto_test.qmd`) and save it in your `scripts/` folder.

You'll see an HTML file appear in your browser. Notice:

-   The first code chunk (`1 + 1`) runs and displays its result.
-   The second chunk shows the result but not the code - that's because
    it includes the code chunk option `#| echo: false`. More on these
    later!

### YAML header

???

### Markdown Syntax

The prose sections of a Quarto document use **Markdown**, a lightweight
formatting language with only \~15 rules to learn. Markdown is also used
in GitHub READMEs, Slack, and many other tools, so it's well worth
knowing.

| Syntax             | Result                 |
|--------------------|------------------------|
| `# Heading 1`      | Large top-level header |
| `## Heading 2`     | Smaller sub-header     |
| `**bold**`         | **bold**               |
| `*italic*`         | *italic*               |
| `[link text](url)` | Hyperlink              |
| `* item`           | Bullet list            |
| `1. item`          | Numbered list          |

For a full reference, see https://www.markdownguide.org/cheat-sheet/.

::: callout-tip
The `#` symbol means something different in R code (a comment) vs.
Markdown (a header). In Markdown, the number of `#` symbols you use
matters: `#` is the biggest header, `###` up to `######` are
consecutively smaller and nested below it.
:::

::: callout-exercise
Add some italic text, a numbered list, and a sub-header to your
document. Try a few other Markdown syntax options. Re-render and check
the output.
:::

### Visual vs. Source Editor

Quarto includes a **Visual Editor** that lets you format text visually, like a
word processor. While we recommend learning raw Markdown (**Source
Editor**) - since you'll use it in many other contexts - the Visual
Editor has two genuinely useful features:

-   **Citations:** Go to **Insert \> Citation**, choose a source (e.g.,
    paste a DOI), and Quarto automatically formats the citation, creates
    a `references.bib` file, and updates your YAML header. Done.
-   **Tables:** Markdown tables are tedious to type by hand. Use the
    **Table** menu to insert and edit them visually.

To toggle between Visual Editor and Source Editor mode:

* RStudio: Click the "Visual" button (vs. "Source") at the top right of the editor pane.
* Positron: Command-Shift-F4 (Mac) or Ctrl-Shift-F4 (Windows).

::: callout-exercise
Enable Visual Editor mode.

Add a citation (e.g., DOI 10.1038/s41467-020-17726-z) into a section of
prose in your Quarto document, using the Visual Editor.

Add a table into your Quarto document using the Visual Editor.

Toggle back to Source Editor mode. Look at the raw Markdown and BibTeX that was generated for your citation and table.

Render the document and check the differences in the output.
:::

### Code Chunks in Quarto

Code chunks contain functional code to be executed.  Markdown syntax will not work here!  A code chunk 
You can create a new chunk in your Quarto in one of these ways:

-   Go to Code in the top menu bar, click "Insert Chunk"
-   Use the keyboard shortcut
    -   RStudio: Mac: Command-Option-i or  Windows: Ctrl-Alt-i
    -   Positron: Mac: Command-Shift-i or  Windows: Ctrl-Shift-i
-   Type by hand:

```{{r}}
```

#### Executing Code Chunks in Interactive Mode

All the R code in your Quarto document will execute when you Render the document, but you can also run it on the fly ("interactive" mode) to test and debug it as you write.

::: {.panel-tabset}
##### RStudio

Multiple options to run code; here are a few:

1.  Select the line (or simply put the cursor there), and click "Run".
    This is available from:
    a.  the bar above the file (green arrow)
    b.  the menu bar: Code \> Run Selected Line(s)
    c.  keyboard shortcut: Command-Enter or Ctrl-Enter
2.  Click the green arrow at the right of the code chunk to run the entire chunk.

##### Positron

Multiple options to run code; here are a few:

1.  Select the line (or simply put the cursor there), and use the keyboard shortcut: Command-Enter or Ctrl-Enter
2.  Click the Run Cell arrow above the code chunk to run the entire chunk.

:::

:::{.callout-exercise}
Add a code chunk to your Quarto document:

```{{r}}
x <- 4 * 8

hights_ft <- c(5.2, 6.0, 5.7)

coef <- 3.14
```

Run the code in interactive mode, and examine the values in the Environment pane (RStudio) or Session Variables pane (Positron).
:::

#### Code chunk options

Here are some common options available to customize code chunk outputs.

| Option              | What it does                                   |
|---------------------|------------------------------------------------|
| `#| eval: false`    | Don't run the code (just display it)           |
| `#| echo: false`    | Run the code, but hide it from the output      |
| `#| message: false` | Suppress diagnostic messages                   |
| `#| warning: false` | Suppress warning messages                      |
| `#| include: false` | Run the code silently - no output shown at all |
| `#| fig-width: 6`   | Set the width (or height) of a plot (in inches)|
| `#| fig-align: center` | Set the alignment of a plot (left, right, center) |

Options can be combined. For example, an exploratory chunk you don't
want in your final report:

```{{r}}
#| eval: false # <1>
#| echo: false # <2>
summary(my_data)
```
1. Don't evaluate/run the code
2. Don't show the code in the rendered output

Many more code chunk options can be found at
[Quarto.org](https://quarto.org/docs/computations/execution-options.html).

## Practice: Literate Analysis with ocean water samples

Now that we have gone over the basics, let's go a little deeper by
building a simple, Quarto document that represents a literate analysis
using real data. We are going to work with the seawater chemistry data.
We are going to use the `BGchem2008data.csv` data we downloaded in our
previous session.

<!-- download a file named `BGchem2008data.csv` from the Arctic Data Center repository. Please follow the steps below to download the data and then upload to your RStudio Server `data` folder. -->

### Getting Started

<!-- ::: callout-tip -->

<!-- #### Set up -->

<!-- 1.  Navigate to this dataset by Craig Tweedie that is published on the Arctic Data Center. [Craig Tweedie. 2009. North Pole Environmental Observatory Bottle Chemistry. Arctic Data Center. doi:10.18739/A25T3FZ8X.](http://doi.org/10.18739/A25T3FZ8X) -->

<!-- 2.  Download the first csv file called `BGchem2008data.csv` by clicking the "download" button next to the file. -->

<!-- 3. Click the "Upload" button in your RStudio server file browser. -->

<!-- 3. In the dialog box, make sure the destination directory is the `data` directory in your R project, click "Choose File," and locate the `BGchem2008data.csv` file. Press "OK" to upload the file. -->

<!-- 4. Check your file was successfully uploaded by navigating into your `data` folder in the **Files** pane. -->

<!-- ::: -->

Experienced R users who have never used Quarto (or RMarkdown) often
struggle a bit in the transition to developing analysis in Prose+Code
format --- which makes sense! It is switching the code paradigm to a new
way of thinking.

Rather than starting an R chunk and putting all of your code in that
single chunk, below we describe what we think is a better way.

1.  Open a document and block out the high-level sections you know
    you'll need to include using top level headers.
2.  Add bullet points for some high level pseudo-code steps you know
    you'll need to take.
3.  Start filling in under each bullet point the code that accomplishes
    each step. As you write your code, transform your bullet points into
    prose, and add new bullet points or sections as needed.

For this mini-analysis, we will have the following sections and code
steps:

1.  Introduction
    a.  About the data
    b.  Setup
    c.  Read in data
2.  Analysis
    a.  Calculate summary statistics
    b.  Calculate mean Redfield ratio
    c.  Plot Redfield ratio
3.  Conclusion

::: callout-note
## Exercise

Under "About the data", write a sentence saying where the data set came
from, including a hyperlink ti the data. Also mention when was the data
downloaded.

Hint: Navigate to Help \> Markdown Quick Reference to look-up the
hyperlink syntax.
:::

### Read in the data

Now that we have outlined our document, we can start writing code! To
read the data into our environment, we will use a function from the
`readr` package.

To use a package in our analysis, we need to first make sure it is
installed (you can install a package by running
`install.package("name-of-package")`). Once installed you need to load
it into our environment using `library(package_name)`. Even though we
have installed it, we haven't yet told our R session to access it.
Because there are so many packages (many with conflicting namespaces) R
cannot automatically load every single package you have installed.
Instead, you load only the ones you need for a particular analysis.
**Loading the package is a key part of the *reproducible* aspect of our
literate analysis**, so we will include it as an R chunk as part of our
Setup.

::: {.callout-caution icon="false"}
## Best Practice

It is generally good practice to include all of your `library()` calls
in a single, dedicated R chunk near the top of your document. This lets
collaborators know what packages they might need to install before they
start running your code.
:::

The server should have already installed the two packages we need for
now: `readr` and `here`. Let's add a new R chunk below your Setup header
that calls these libraries, and run it.

It should look like this:

```{r}
#| eval: false

library(readr)
library(here)
```

::: {.callout-caution icon="false"}
## Quarto file path and the `here()` function

Quarto has a special way of handling relative paths that can be very
handy. When working in an Quarto document, **R will set all paths
relative to the location of the Quarto file**. This can make things
easier to read in data if your Quarto document is stored in the same
directory or "near" by. However, more often that not, your `.qmd` file
will be stored in a a folder (e.g `scripts`) and your data in a `data`
folder, (both folder in the main project directory).

The `here()` function helps navigate this file path mix up in a straight
forward and reproducible way. This function sets the file path to the
project's directory and builds the rest of the file path from there.
Making it easier to find files inside different folders in a project. In
this case, because the `.qmd` file lives in the script folder, `here()`
makes is easy to navigate back into the project's directory and then
into the `data` folder to read in our file.
:::

Now, under "Read data", add a code chunk that uses the `read_csv()` with
the `here()` function to read in your data file.

```{r load data for quarto book}
#| echo: false
library(readr)

### Quarto book structure: script-specific data in script-specific folder
bg_chem <- read_csv("data/r_quarto_literate_analysis/BGchem2008data.csv")
```

```{r demonstrate how to load data from student repo}
#| eval: false

bg_chem <- read_csv(here::here("data/BGchem2008data.csv"))
```

::: column-margin
**Why `read_csv()` over `read.csv()`?**

We chose to show `read_csv()` from the `readr` package to introduce the
concept of packages, to show you how to load packages, and `read_csv()`
has several advantages over `read.csv()` from base R, including:

-   More reasonable function defaults (no `stringsAsFactors`!)
-   Smarter column type parsing, especially for dates
-   `read_csv()` is much faster than `read.csv()`, which is helpful for
    large files
:::

Once you run this line in your document, you should see the `bg_chem`
object populate in your environment pane. It also spits out lots of text
explaining what types the function parsed each column into. This text is
important, and should be examined, but we might not want it in our final
document.

::: callout-note
## Exercise

How would you suppress the warnings (so they don't show in our output
file) form a specific code chunk?

**Hint:** Code chunk options
:::

### Calculate Summary Statistics

As our "analysis" we are going to calculate some very simple summary
statistics and generate a single plot. Using water samples from the
Arctic Ocean, we will examine the ratio of nitrogen to phosphate to see
how closely the data match the Redfield ratio, which is the consistent
16:1 ratio of nitrogen to phosphorous atoms found in marine
phytoplankton.

Let's start by exploring the data we just read. Every time we read a new
data set, it is important to familiarize yourself with it and make sure
that the data looks as expected. Below some useful functions for
exploring your data.

Let's start by creating a new R chunk and run the following functions.
Because this just an exploration and we do not want this chunk to be
part of our report, we will indicate that by adding `#|eval: false` and
`#| echo: false` in the setup of the chunk, that way, the code in this
chunk will not run and not be displayed when I knit the final document.

```{r}
#| eval: false

## Prints the column names of my data frame
colnames(bg_chem)

## General structure of the data frame - shows class of each column
str(bg_chem)

## First 6 lines of the data frame
head(bg_chem)

## Summary of each column of data
summary(bg_chem)

## Prints unique values in a column (in this case Date)
unique(bg_chem$Date)
```

To peek out data frame, we can type `View(bg_chem)` in the console. This
will open a tab with our data frame in a tabular format.

Now that we know a more about the data set we are working with lets do
some analyses. Under the appropriate bullet point in your analysis
section, create a new R chunk, and use it to calculate the mean nitrate
(NO3), nitrite (NO2), ammonium (NH4), and phosphorous (P) measured.

Save these mean values as new variables with easily understandable
names, and write a (brief) description of your operation using markdown
above the chunk. Remember that the `$` (aka the subset operator)
indicates which column of your data to look into.

```{r}
nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)
```

In another chunk, use those variables to calculate the nitrogen:
phosphate ratio (Redfield ratio).

```{r}
ratio <- (nitrate + nitrite + amm)/phos
```

You can access this variable in your markdown text by using R in-line in
your text. The syntax to call R in-line (as opposed to as a chunk) is a
single back tick \`, followed by the letter "r", then whatever your
simple R command is --- here we will use `round(ratio)` to print the
calculated ratio, and finally a closing back tick \`. This allows us to
access the value stored in this variable in our explanatory text without
resorting to the evaluate-copy-paste method so commonly used for this
type of task.

So, the text in you Quarto document should look like this:

The Redfield ratio for this dataset is approximately: \`r
`round(ratio)`\`

And the rendered text like this:

The Redfield ratio for this dataset is approximately `r round(ratio)`.

Finally, create a simple plot using base R that plots the ratio of the
individual measurements, as opposed to looking at mean ratio.

```{r}
plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)
```

::: callout-note
## Exercise

Decide whether or not you want the plotting code above to show up in
your knitted document along with the plot, and implement your decision
as a chunk option.

Render your Quarto document (by pressing the Render button) and observe
the results.
:::

::: callout-important
## How do I decide when to make a new code chunk?

Like many of life's great questions, there is no clear cut answer. A
rule of thumb is to have one chunk per functional unit of analysis. This
functional unit could be 50 lines of code or it could be 1 line, but
typically it only does one "thing." This could be reading in data,
making a plot, or defining a function. It could also mean calculating a
series of related summary statistics (as we'll see below). Ultimately,
the choice is one related to personal preference and style, but
generally you should ensure that code is divided up such that it is
easily explainable in a literate analysis as the code is run.
:::

## Quarto and Environments

Let's walk through an exercise with the document we just created to
demonstrate how Quarto handles environments. We will be deliberately
inducing some errors here for demonstration purposes.

First, follow these steps:

::: callout-tip
## Setup

-   Restart your R session (Session \> Restart R)
-   Run the last chunk in your Quarto document by pressing the play
    button on the chunk
:::

Perhaps not surprisingly, we get an error:

```         
Error in plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4) : 
  object 'bg_chem' not found
```

This is because we have not run the chunk of code that reads in the
`bg_chem` data. The R part of Quarto works just like a regular R script.
You have to execute the code, and the order that you run it in matters.
It is relatively easy to get mixed up in a large Quarto document ---
running chunks out of order, or forgetting to run chunks.

To resolve this, follow the next step:

::: callout-tip
## Setup continued

-   Select from the "Run" menu (top right of the editor pane) "Run All."
-   Observe the `bg_chem` variable in your environment
:::

This is a great way to reset and re-run code when things seem to have
gone sideways. It is great practice to do periodically since it helps
ensure you are writing code that actually runs and it's reproducible.

::: callout-tip
## For the next exercise:

-   Clean your environment by clicking the broom in the environment pane
-   Restart your R session (Session \> Restart R)
-   Press "Render" to run all of the code in your document
-   Observe the state of your environment pane

Assuming your document rendered and produced an html page, your code
ran. Yet, the environment pane is empty. What happened?
:::

The Render button is rather special --- it doesn't just run all of the
code in your document. It actually spins up a fresh R environment
separate from the one you have been working in, runs all of the code in
your document, generates the output, and then closes the environment.
This is one of the best ways Quarto (or RMarkdown) helps ensure you have
built a reproducible workflow. If, while you were developing your code,
you ran a line in the console as opposed to adding it to your Quarto
document, the code you develop while working actively in your
environment will still work. However, when you knit your document, the
environment RStudio spins up doesn't know anything about that working
environment you were in. Thus, your code may error because it doesn't
have that extra piece of information. Commonly, `library()` calls are
the source of this kind of frustration when the author runs it in the
console, but forgets to add it to the script.

To further clarify the point on environments, perform the following
steps:

::: callout-tip
## Setup continued

-   Select from the "Run" menu (top right of editor pane) "Run All"
-   Observe all of the variables in your environment
:::

::: callout-important
## What about all my R Scripts?

Some pieces of R code are better suited for R scripts than Quarto or
RMarkdown. A function you wrote yourself that you use in many different
analyses is probably better to define in an R script than repeated
across many Quarto or RMarkdown documents. Some analyses have mundane or
repetitive tasks that don't need to be explained very much. For example,
in the document shown in the beginning of this lesson, 15 different
excel files needed to be reformatted in slightly different, mundane
ways, like renaming columns and removing header text. Instead of
including these tasks in the primary Quarto document, the authors chose
to write one R script per file and stored them all in a directory. Then,
took the contents of one script and included it in the literate
analysis, using it as an example to explain what the scripts did, and
then used the `source()` function to run them all from within the Quarto
document.

So, just because you know Quarto now, doesn't mean you won't be using R
scripts anymore. Both `.R` and `.qmd` have their roles to play in
analysis. With practice, it will become more clear what works well in
Quarto or RMarkdown, and what belongs in a regular R script.
:::

## Additional Quarto Resources

-   Posit (the organization that developed Quarto) has great
    documentation, check out [Quarto.org](https://quarto.org/)
-   R for Data Science (2e) (Wickham et al, 2023), this is an awesome
    book for all R related things. Chapter [29 and
    30](https://r4ds.hadley.nz/quarto.html) are specific to Quarto.
-   [Quarto Gallery:](https://quarto.org/docs/gallery/) Example of
    different outputs created using Quarto
-   [Hello Quarto: share, collaborate, teach,
    reimagine](https://openscapes.org/blog/2022-08-10-quarto-keynote/).
    A talk by Julia Stewart Lowndes and Mine Cetinkaya-Runde.

## Troubleshooting: My RMarkdown/Quarto doc Won't Knit to PDF

If you get an error when trying to knit to PDF that says your computer
doesn't have a LaTeX installation, one of two things is likely
happening:

-   Your computer doesn't have LaTeX installed
-   You have an installation of LaTeX but RStudio cannot find it (it is
    not on the path)

If you already use LaTeX (like to write papers), you fall in the second
category. Solving this requires directing RStudio to your installation -
and isn't covered here.

If you fall in the first category - you are sure you don't have LaTeX
installed - can use the R package `tinytex` to easily get an
installation recognized by RStudio, as long as you have administrative
rights to your computer.

To install `tinytex` run:

```{r}
#| eval: false
install.packages("tinytex")
tinytex::install_tinytex()
```

If you get an error that looks like destination /usr/local/bin not
writable, you need to give yourself permission to write to this
directory (again, only possible if you have administrative rights). To
do this, run this command in the terminal:

``` bash
sudo chown -R `whoami`:admin /usr/local/bin
```

and then try the above install instructions again. Learn more about
`tinytex` from [Yihui Xie's online book
*TinyTeX*](https://yihui.org/tinytex/). \`\`\`\`
